---
title: "P8105_HW3_sx2289"
author: "Shuhong Xinag"
date: "10/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
devtools::install_github("p8105/p8105.datasets")
```
```{r}
library(tidyverse)
```


##Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user/order variables -- user ID, order ID, order day, and order hour .There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```

##Problem 2

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accelerometer = 
read_csv("./data/accel_data.csv") %>%
  janitor::clean_names()%>%
  pivot_longer(
    cols = starts_with("activity"),
    names_to = "min",
    names_prefix = "activity_",
    values_to = "activity",
    values_drop_na = TRUE)%>%
  mutate(
    weekend_vs_weekday = case_when(
      day %in% c("Saturday", "Sunday") ~"weenkend",
      !(day %in% c("Saturday","Sunday")) ~ "weekday" 
    )
  ) %>%
  mutate(
    weekend_vs_weekday = as.factor(weekend_vs_weekday),
    week = as.factor(week),
    day_id = as.factor(day_id), 
    day = as.factor(day),
    min = as.numeric(min),
    activity = as.numeric(activity)
  ) %>%
  mutate(day = forcats::fct_relevel(day,
                                   c("Monday","Tuesday","Wednesday","Thursday","Friday", "Saturday","Sunday")))%>%
  group_by(week)%>%
  arrange(day, .by_group = TRUE)

accelerometer
```

This final dataset contains `r nrow(accelerometer)` rows and `r ncol(accelerometer)` columns. Observations in the experiments are the activity counts for each minute of a 24-hour day starting at midnight for five week. There are 1440 activity counts in total. 

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accelerometer %>%
  group_by(week,day)%>%
  summarise( total_activity_each_day = sum(activity, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = day,
    values_from = total_activity_each_day
  )%>%
knitr::kable()
```

From the table we can see this individual does bot have a stable activity habit. There is no apparent tends. But we can see n Wednesday, the person has a relatively stable activity but on Sunday, the activity counts each minute are keeping decreasing. 

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accelerometer %>% 
  group_by(day_id,day, week) %>% 
  ggplot(aes(x = min,
             y = activity,
             color = day,
             group = day_id))+
  geom_point()+
  stat_smooth(
    se = F,
    method = "loess",
    alpha = 0.5
  )+
  scale_y_continuous(name = "Hourly Activity Counts", 
                      trans = "sqrt")+
   scale_x_continuous(name ="Hours" ,
                     breaks = c(seq(120,1440,120)), 
                     labels = c("2am", "4am", "6am", "8am", "10am", "12pm","2pm", "4pm", "6pm","8pm", "10pm", "12am")) +
  viridis::scale_color_viridis(discrete = T,
                               name = "")+
  labs(title = "The 24-Hour activities in 5 Weeks for the 63 year-old male with BMI 25 ")
ggsave("Problem 2_Hourly Activity.jpg")
```

From the graph, we can tell that 63 year-old male with BMI 25 usually does activities during the time from around 8 am to 10 pm. Also, he less active during the weekends.

##Problem 3

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r}
tidy_ny_noaa = 
  ny_noaa %>%
  separate(date, into = c("year","month","day")) %>%
  mutate_at(vars(year, month, day), as.factor) %>%
  mutate_at(vars(prcp, snow, tmax, tmin), as.numeric) %>%
  mutate(
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10,
  )
  
skimr::skim_without_charts(tidy_ny_noaa)
```

```{r}
tidy_ny_noaa %>%
  filter(snow != "NA") %>%
  count(snow)%>%
  filter(snow >= "0")%>%
  arrange(desc(n))
```

The most commonly value is 0, which means most of time there is no snow in NY. 


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
plot_Jan_July = 
  tidy_ny_noaa %>%
  filter(month %in% c("01","07")) %>% 
  group_by(year, month, id) %>% 
  summarize(
    mean_tmax = mean(tmax),na.rm = T) %>%
  drop_na() %>%
  ggplot(
    aes(
      x = year,
      y = mean_tmax,
      color = id,
      group = id))+
geom_point()+
geom_path()+
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "Highest mean temperature in January and in July",
       x = "Year",
       y = "Mean Temperature(C)")

ggsave("plot_Jan_July.jpg")
```

From the graph, we can learned that the temperature keeps relatively stable throughout the 30 years. There was only 2 times that the lowest temperature were below the -10 C and the highest temperature were around the 30 C. For the out liner, we can see some of them fro the graph, but the number of them was very small.

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

plot_i = 

